---
title: "Congress experience"
author: "Rebecca S. Chen & Tuba Rizvi"
---

```{r setup, echo=F, message=F}

# load packages
pacman::p_load(lme4, tidyverse, data.table, kableExtra, lmtest, MASS)

### Load data ####

#survey
load("../data/post_survey/clean_survey_all_countrycat_add.RData")
survey$gender <- gsub("Prefer_not_say", NA, survey$gender)
survey$lgbtq <- gsub("Prefer not to say", NA, survey$lgbtq)

### Set theme ####
source("../scripts/EDI_ggplot_theme.R")

### Load function to output model results
source("../scripts/function_collect_model_output.R")
```

**How did people with different social identities experience the congress?**

## Discrimination and harassment

Respondents of the post-congress survey were asked if they themselves experienced discrimination and/or harassment (of any sort) at the congress and whether they reported it to the Awareness team, or if they witnessed someone else experiencing this.

A total of 11 respondents experienced some form of discrimination or harassment, of which 2 cases were reported to the Awareness team. A total of 3 respondents witnessed somebody else receiving some form of discrimination or harassment, of which 1 case was reported to the Awareness team. Reasons of not reporting the cases to the Awareness team included that the case was unrelated to EDI issues. Even though only a few cases were reported to the Awareness team, the qualitative feedback given in the survey highlighted that the presence of this team made some people feel safe.

Due to the low number of reports of discrimination or harassment, we cannot statistically test if some social identities experienced more discrimination/harassment than others. Nevertheless, these are the summary statistics per social identity for the answer to whether respondents experienced discrimination/harassment themselves:

```{r discriminate, echo = T, warning=F, message =F}
dis <- subset(survey, discrimination == "Yes")

table(dis$gender, dis$lgbtq, dis$nationality_subcontinent)
```

This shows that most reports came from women, and from West Europe.

## Congress experience

In the post-congress survey, we asked respondents to answer on a 7-point Likert scale (1: Strongly disagree, 7: Strongly agree) how much they agree with the following three statements: 1) "I felt heard during the conversations I had, both during Q&A sessions and social activities"; 2) "I felt comfortable being myself"; 3) "Attending the Behaviour 2023 congress helped me feel like I belong in my research field". For each of the three statements, we fitted ordinal GLMs to identify which social identity variables (gender, LGBTQIA+, nationality, affiliation, expat status) were significantly associated with the Likert-scale response to the statement. Additionally, we controlled for the level of comfort a person had speaking English as well as their self-reported level of expertise. Only the variables that were significant were used in the final model.

We have information on three variables that roughly measure the same thing: age is indicative of career stage, career stage is indicative of expertise, and self-reported expertise rating directly measures this expertise. Since age and career stage are indicator variables of expertise, we therefore only investigate expertise rating as this will give us the signal we are interested in.

### Feeling heard

```{r feeling heard, echo=T, warning=F, message=F}
# first, we explore the distribution of answers

# feeling heard
ggplot(survey, aes(feeling_heard_rating)) + 
  geom_histogram(stat="count", aes(y=stat(count/sum(count)*100)), fill = clrs[11]) + 
  geom_text(aes(label = paste0("N = ", ..count..), y = stat(count/sum(count)*100)), 
            stat="count", vjust=-1) +
  ylim(0, 100) + 
  labs(x = "I felt heard", y = "% of responses") 

# next: explore the relationship between age and expertise, and career stage and expertise rating

coeftest(polr(as.factor(expertise_rating) ~ age, data = survey))
coeftest(polr(as.factor(expertise_rating) ~ career_3cat, data = survey))

# as expected!

# another check: is expertise rating affected by gender?

coeftest(polr(as.factor(expertise_rating) ~ gender, data = survey))

# answer = no

# then we test the effect of each of the social identity variables

# gender
survey$gender <- factor(survey$gender, levels = c("Male", "Female", "Non-binary"))

m_heard_gender_null <- polr(feeling_heard_rating ~ 1, data=subset(survey, !is.na(gender)))

m_heard_gender <- polr(feeling_heard_rating ~ gender,
                   data=survey) 

m_heard_gender_out <- collect_out(model = m_heard_gender, null = m_heard_gender_null, name = "feeling_heard_gender", n_factors = 2, type = "likert", save = "yes",  dir = "../results/survey") 

m_heard_gender_out %>% t() %>% kbl() %>%
  kable_classic_2() # not significant

# lgbtqia 

m_heard_lgbtq_null <- polr(feeling_heard_rating ~ 1, data=subset(survey, !is.na(lgbtq)))

m_heard_lgbtq <- polr(feeling_heard_rating ~ lgbtq,
                   data=subset(survey, !is.na(lgbtq))) 

m_heard_lgbtq_out <- collect_out(model = m_heard_lgbtq, null = m_heard_lgbtq_null, name = "feeling_heard_lgbtq", n_factors = 1, type = "likert", save = "yes",  dir = "../results/survey") 

m_heard_lgbtq_out %>% t() %>% kbl() %>%
  kable_classic_2() # almost significant, note that the baseline is answer "No" to lgbtq identity

# nationality
survey$nationality_continent <- factor(survey$nationality_continent, levels = c("Europe", "Asia", "North America", "Oceania", "South America"))

m_heard_nat_null <- polr(feeling_heard_rating ~ 1, data=subset(survey, !is.na(nationality_continent)))

m_heard_nat <- polr(feeling_heard_rating ~ nationality_continent,
                   data=survey) 

m_heard_nat_out <- collect_out(model = m_heard_nat, null = m_heard_nat_null, name = "feeling_heard_nat", n_factors = 4, type = "likert", save = "yes", 
            dir = "../results/survey") 

m_heard_nat_out %>% t() %>% kbl() %>%
  kable_classic_2() # almost significant, but Asia is borderline sig with lower values

# affiliation
survey$affiliation_continent <- factor(survey$affiliation_continent, levels = c("Europe", "Asia", "Africa", "North America", "Oceania", "South America"))

m_heard_aff_null <- polr(feeling_heard_rating ~ 1, data=subset(survey, !is.na(affiliation_continent)))

m_heard_aff <- polr(feeling_heard_rating ~ affiliation_continent,
                   data=subset(survey, !is.na(affiliation_continent))) 

m_heard_aff_out <- collect_out(model = m_heard_aff, null = m_heard_aff_null, name = "feeling_heard_aff", n_factors = 5, type = "likert", save = "yes", dir = "../results/survey") 

m_heard_aff_out %>% t() %>% kbl() %>%
  kable_classic_2() # not significant

# expat

survey$expat <- factor(survey$expat, levels = c("No expat", "Expat"))

m_heard_expat_null <- polr(feeling_heard_rating ~ 1, data=subset(survey, !is.na(expat)))

m_heard_expat <- polr(feeling_heard_rating ~ expat,
                   data=subset(survey, !is.na(expat))) 

m_heard_expat_out <- collect_out(model = m_heard_expat, null = m_heard_expat_null, name = "feeling_heard_expat", n_factors = 1, type = "likert", save = "yes",  dir = "../results/survey") 

m_heard_expat_out %>% t() %>% kbl() %>%
  kable_classic_2() # not significant

# english level 
survey$english_comfort_rating <- as.numeric(as.character(survey$english_comfort_rating))

m_heard_english_null <- polr(feeling_heard_rating ~ 1, data=subset(survey, !is.na(english_comfort_rating)))

m_heard_english <- polr(feeling_heard_rating ~ english_comfort_rating,
                   data=subset(survey, !is.na(english_comfort_rating))) 

m_heard_english_out <- collect_out(model = m_heard_english, null = m_heard_english_null, name = "feeling_heard_english", n_factors = 1, type = "likert", save = "yes",  dir = "../results/survey") 

m_heard_english_out %>% t() %>% kbl() %>%
  kable_classic_2() # significant

# expert rating
survey$expertise_rating <- as.numeric(as.character(survey$expertise_rating))

m_heard_expert_null <- polr(feeling_heard_rating ~ 1, data=subset(survey, !is.na(expertise_rating)))

m_heard_expert <- polr(feeling_heard_rating ~ expertise_rating,
                   data=subset(survey, !is.na(expertise_rating))) 

m_heard_expert_out <- collect_out(model = m_heard_expert, null = m_heard_expert_null, name = "feeling_heard_expert", n_factors = 1, type = "likert", save = "yes",  dir = "../results/survey") 

m_heard_expert_out %>% t() %>% kbl() %>%
  kable_classic_2() # significant

### build final model only with significant variables

m_heard_null <- polr(feeling_heard_rating ~ 1, data=subset(survey, !is.na(expertise_rating) & !is.na(english_comfort_rating)))

m_heard <- polr(feeling_heard_rating ~english_comfort_rating + expertise_rating, data=subset(survey, !is.na(expertise_rating) & !is.na(english_comfort_rating)))

drop1(m_heard, test = "Chisq")

coeftest(m_heard)

m_heard_out <- collect_out(model = m_heard, null = m_heard_null, name = "feeling_heard_final", n_factors = 2, type = "likert", save = "yes",  dir = "../results/survey") 

m_heard_out %>% t() %>% kbl() %>%
  kable_classic_2() 

```

The results indicate that none of the social identity variables (gender, lgbtq, nationality, affiliation (but this model did not converge)) affected whether a person felt heard during the congress. However, people who are more comfortable speaking English felt heard more, and people that rated their expertise in the field higher also felt heard more.

### Feeling comfortable being yourself

```{r being yourself, echo=T, warning=F, message=F}
# first, we explore the distribution of answers

ggplot(survey, aes(comfort_being_yourself_rating)) + 
  geom_histogram(stat="count", aes(y=stat(count/sum(count)*100)), fill = clrs[11]) + 
  geom_text(aes(label = paste0("N = ", ..count..), y = stat(count/sum(count)*100)), 
            stat="count", vjust=-1) +
  ylim(0, 100) + 
  labs(x = "I felt comfortable being myself", y = "% of responses") 

# then we test the effect of each of the social identity variables

# gender
m_yourself_gender_null <- polr(comfort_being_yourself_rating ~ 1, data=subset(survey, !is.na(gender)))

m_yourself_gender <- polr(comfort_being_yourself_rating ~ gender,
                   data=survey) 

m_yourself_gender_out <- collect_out(model = m_yourself_gender, null = m_yourself_gender_null, name = "comf_yourself_gender", n_factors = 2, type = "likert", save = "yes",  dir = "../results/survey") 

m_yourself_gender_out %>% t() %>% kbl() %>%
  kable_classic_2() # significant

# lgbtqia 

m_yourself_lgbtq_null <- polr(comfort_being_yourself_rating ~ 1, data=subset(survey, !is.na(lgbtq)))

m_yourself_lgbtq <- polr(comfort_being_yourself_rating ~ lgbtq,
                   data=subset(survey, !is.na(lgbtq))) 

m_yourself_lgbtq_out <- collect_out(model = m_yourself_lgbtq, null = m_yourself_lgbtq_null, name = "comf_yourself_lgbtq", n_factors = 1, type = "likert", save = "yes",  dir = "../results/survey") 

m_yourself_lgbtq_out %>% t() %>% kbl() %>%
  kable_classic_2() # almost significant

# nationality
m_yourself_nat_null <- polr(comfort_being_yourself_rating ~ 1, data=subset(survey, !is.na(nationality_continent)))

m_yourself_nat <- polr(comfort_being_yourself_rating ~ nationality_continent,
                   data=survey) 

m_yourself_nat_out <- collect_out(model = m_yourself_nat, null = m_yourself_nat_null, name = "comf_yourself_nat", n_factors = 4, type = "likert", save = "yes", 
            dir = "../results/survey") 

m_yourself_nat_out %>% t() %>% kbl() %>%
  kable_classic_2() # not significant, but East Europe is sig with higher values

# affiliation
m_yourself_aff_null <- polr(comfort_being_yourself_rating ~ 1, data=subset(survey, !is.na(affiliation_continent)))

m_yourself_aff <- polr(comfort_being_yourself_rating ~ affiliation_continent,
                   data=subset(survey, !is.na(affiliation_continent))) 

m_yourself_aff_out <- collect_out(model = m_yourself_aff, null = m_yourself_aff_null, name = "comf_yourself_aff", n_factors = 5, type = "likert", save = "yes", dir = "../results/survey") 

m_yourself_aff_out %>% t() %>% kbl() %>%
  kable_classic_2() # not significant

# expat

m_yourself_expat_null <- polr(comfort_being_yourself_rating ~ 1, data=subset(survey, !is.na(expat)))

m_yourself_expat <- polr(comfort_being_yourself_rating ~ expat,
                   data=subset(survey, !is.na(expat))) 

m_yourself_expat_out <- collect_out(model = m_yourself_expat, null = m_yourself_expat_null, name = "comf_yourself_expat", n_factors = 1, type = "likert", save = "yes",  dir = "../results/survey") 

m_yourself_expat_out %>% t() %>% kbl() %>%
  kable_classic_2() # not significant

# english level 
survey$english_comfort_rating <- as.numeric(as.character(survey$english_comfort_rating))

m_yourself_english_null <- polr(comfort_being_yourself_rating ~ 1, data=subset(survey, !is.na(english_comfort_rating)))

m_yourself_english <- polr(comfort_being_yourself_rating ~ english_comfort_rating,
                   data=subset(survey, !is.na(english_comfort_rating))) 

m_yourself_english_out <- collect_out(model = m_yourself_english, null = m_yourself_english_null, name = "comf_yourself_english", n_factors = 1, type = "likert", save = "yes",  dir = "../results/survey") 

m_yourself_english_out %>% t() %>% kbl() %>%
  kable_classic_2() # significant

# expert rating
survey$expertise_rating <- as.numeric(as.character(survey$expertise_rating))

m_yourself_expert_null <- polr(comfort_being_yourself_rating ~ 1, data=subset(survey, !is.na(expertise_rating)))

m_yourself_expert <- polr(comfort_being_yourself_rating ~ expertise_rating,
                   data=subset(survey, !is.na(expertise_rating))) 

m_yourself_expert_out <- collect_out(model = m_yourself_expert, null = m_yourself_expert_null, name = "comf_yourself_expert", n_factors = 1, type = "likert", save = "yes",  dir = "../results/survey") 

m_yourself_expert_out %>% t() %>% kbl() %>%
  kable_classic_2() # significant

### build final model only with significant variables

m_yourself_null <- polr(comfort_being_yourself_rating ~ 1, data=subset(survey, !is.na(gender) & !is.na(expertise_rating) & !is.na(english_comfort_rating)))

m_yourself <- polr(comfort_being_yourself_rating ~ gender + english_comfort_rating + expertise_rating, data=subset(survey, !is.na(gender) & !is.na(expertise_rating) & !is.na(english_comfort_rating)))

drop1(m_yourself, test = "Chisq")

m_yourself_out <- collect_out(model = m_yourself, null = m_yourself_null, name = "comf_yourself_final", n_factors = 4, type = "likert", save = "yes",  dir = "../results/survey") 

m_yourself_out %>% kbl() %>%
  kable_classic_2() 

```

The results indicate that lgbtq and nationality did not affect whether a person felt comfortable being themselves during the congress (but lgbtq came close to significance). However, men feel more comfortable than women and non-binary people feel less comfortable being themselves compared to women. People who are more comfortable speaking English felt more comfortable being themselves, and people that rated their expertise in the field higher also felt more comfortable being themselves.

### Sense of belonging

```{r sense of belonging, echo=T, warning=F, message=F}
# first, we explore the distribution of answers

ggplot(survey, aes(sense_of_belonging_rating)) + 
  geom_histogram(stat="count", aes(y=stat(count/sum(count)*100)), fill = clrs[11]) + 
  geom_text(aes(label = paste0("N = ", ..count..), y = stat(count/sum(count)*100)), 
            stat="count", vjust=-1) +
  ylim(0, 100) + 
  labs(x = "Attending the Behaviour 2023 congress helped me 
feel like I belong in my research field", y = "% of responses") 

# then we test the effect of each of the social identity variables

# gender
m_sob_gender_null <- polr(sense_of_belonging_rating ~ 1, data=subset(survey, !is.na(gender)))

m_sob_gender <- polr(sense_of_belonging_rating ~ gender,
                   data=survey) 

m_sob_gender_out <- collect_out(model = m_sob_gender, null = m_sob_gender_null, name = "belonging_gender", n_factors = 2, type = "likert", save = "yes",  dir = "../results/survey") 

m_sob_gender_out %>% t() %>% kbl() %>%
  kable_classic_2() # not significant


# lgbtqia 

m_sob_lgbtq_null <- polr(sense_of_belonging_rating ~ 1, data=subset(survey, !is.na(lgbtq)))

m_sob_lgbtq <- polr(sense_of_belonging_rating ~ lgbtq,
                   data=subset(survey, !is.na(lgbtq))) 

m_sob_lgbtq_out <- collect_out(model = m_sob_lgbtq, null = m_sob_lgbtq_null, name = "belonging_lgbtq", n_factors = 1, type = "likert", save = "yes",  dir = "../results/survey") 

m_sob_lgbtq_out %>% t() %>% kbl() %>%
  kable_classic_2() # not significant

# nationality
m_sob_nat_null <- polr(sense_of_belonging_rating ~ 1, data=subset(survey, !is.na(nationality_continent)))

m_sob_nat <- polr(sense_of_belonging_rating ~ nationality_continent,
                   data=survey) 

m_sob_nat_out <- collect_out(model = m_sob_nat, null = m_sob_nat_null, name = "belonging_nat", n_factors = 4, type = "likert", save = "yes", 
            dir = "../results/survey") 

m_sob_nat_out %>% t() %>% kbl() %>%
  kable_classic_2() # not significant, but North America is borderline sig with higher values

# affiliation
m_sob_aff_null <- polr(sense_of_belonging_rating ~ 1, data=subset(survey, !is.na(affiliation_continent)))

m_sob_aff <- polr(sense_of_belonging_rating ~ affiliation_continent,
                   data=subset(survey, !is.na(affiliation_continent))) 

m_sob_aff_out <- collect_out(model = m_sob_aff, null = m_sob_aff_null, name = "belonging_aff", n_factors = 5, type = "likert", save = "yes", 
             dir = "../results/survey") 

m_sob_aff_out %>% t() %>% kbl() %>%
  kable_classic_2() # significant

# expat

m_sob_expat_null <- polr(sense_of_belonging_rating ~ 1, data=subset(survey, !is.na(expat)))

m_sob_expat <- polr(sense_of_belonging_rating ~ expat,
                   data=subset(survey, !is.na(expat))) 

m_sob_expat_out <- collect_out(model = m_sob_expat, null = m_sob_expat_null, name = "belonging_expat", n_factors = 1, type = "likert", save = "yes",  dir = "../results/survey") 

m_sob_expat_out %>% t() %>% kbl() %>%
  kable_classic_2() # not significant

# english level 
survey$english_comfort_rating <- as.numeric(as.character(survey$english_comfort_rating))

m_sob_english_null <- polr(sense_of_belonging_rating ~ 1, data=subset(survey, !is.na(english_comfort_rating)))

m_sob_english <- polr(sense_of_belonging_rating ~ english_comfort_rating,
                   data=subset(survey, !is.na(english_comfort_rating))) 

m_sob_english_out <- collect_out(model = m_sob_english, null = m_sob_english_null, name = "belonging_english", n_factors = 1, type = "likert", save = "yes",  dir = "../results/survey") 

m_sob_english_out %>% t() %>% kbl() %>%
  kable_classic_2() # significant

# expert rating
survey$expertise_rating <- as.numeric(as.character(survey$expertise_rating))

m_sob_expert_null <- polr(sense_of_belonging_rating ~ 1, data=subset(survey, !is.na(expertise_rating)))

m_sob_expert <- polr(sense_of_belonging_rating ~ expertise_rating,
                   data=subset(survey, !is.na(expertise_rating))) 

m_sob_expert_out <- collect_out(model = m_sob_expert, null = m_sob_expert_null, name = "belonging_expert", n_factors = 1, type = "likert", save = "yes",  dir = "../results/survey") 

m_sob_expert_out %>% t() %>% kbl() %>%
  kable_classic_2() # significant

### build final model only with significant variables

m_sob_null <- polr(sense_of_belonging_rating ~ 1, data=subset(survey, !is.na(expertise_rating) & !is.na(english_comfort_rating) & !is.na(affiliation_continent)))

m_sob <- polr(sense_of_belonging_rating ~ affiliation_continent + english_comfort_rating + expertise_rating, data=subset(survey, !is.na(expertise_rating) & !is.na(english_comfort_rating)& !is.na(affiliation_continent)))

drop1(m_sob, test = "Chisq")

coeftest(m_sob)

m_sob_out <- collect_out(model = m_sob, null = m_sob_null, name = "belonging_final", n_factors = 7, type = "likert", save = "yes",  dir = "../results/survey") 

m_sob_out %>% t() %>% kbl() %>%
  kable_classic_2() 

```

The results are very similar to the "Feeling heard" part: none of the social identity variables (gender, lgbtq, nationality) affected whether a person felt that attending the congress increased their Sense of Belonging. However, people who are more comfortable speaking English felt more like it did, and people that rated their expertise in the field higher also felt like attending the congress increased their feeling like they belong in the field.
