---
title: "What other gender disparities do we observe in oral sessions?"
author: "Rebecca S. Chen & Petroula Botsidou"
---

```{r setup, echo=F, message=F}

# load packages
pacman::p_load(lme4, tidyverse, data.table, kableExtra, lmtest)

### Load data ####

#behavioural
load("../data/question_asking/question_asking_data_condensed_for_analysis.RData")

### Set theme ####
source("../scripts/EDI_ggplot_theme.R")

### Load function to output model results
source("../scripts/function_collect_model_output.R")

# subset data

data_analysis <- subset(data_analysis, is.na(followup) & 
                     is.na(host_asks), !is.na(question_type_e)& 
                    !grepl("speaker|questioner", allocator_question))

### convert giving compliments ####

data_analysis$compliment <- gsub("N", "No", data_analysis$compliment)
data_analysis$compliment <- gsub("Y", "Yes", data_analysis$compliment)
data_analysis$compliment  <- as.factor(data_analysis$compliment )

```

*Note: this analysis is only part of the supplementary information of the manuscript*

**What other gender disparities can we observe in oral sessions?**

More specifically, we also collected data on whether positive appraisal was given by a questioner, whether a person asked a question without raising their hand or being chosen by the session host to do so (a jumper), whether a speaker talked longer than their allocated time, and if a question was criticizing the speaker. Depending on the question, we investigated the effect of the question asker's gender, speaker's gender, host's gender while correcting for appropriate confounders such as age or talk number.

For all analysis, we only excluded questions that were follow-up questions, and those asked by the host. This means that these data include manipulated and unmanipulated sessions, as we do not expect our manipulation to interfere with what a questioner says exactly.

## Positive appraisal

Positive appraisal was defined as any positive words towards the speaker, but excluding when simply "Thank you for your talk" was said. However, if this comment included a compliment such as "Thank you for your *nice* talk", this was counted as positive appraisal.

We used binomial GLMMs to address the effect of the gender and age of the questioner and gender and career stage of the speaker, and corrected for the question number within that Q&A session. We assessed the fit of each variable using a likelihood ratio test and included only the variables that explained significant variation in the final model.

### Receiving

```{r compliment_receive, echo=T, warning=F, message=F}

# explore the data 
table(data_analysis$compliment, data_analysis$questioner_gender) %>% kbl() %>%
  kable_classic_2()

table(data_analysis$compliment, data_analysis$speaker_gender) %>% kbl() %>%
  kable_classic_2()

# build initial models

## speaker gender

model_compliment_s_gender <- glmer(compliment ~ speaker_gender + (1|session_id/talk_id), data = subset(data_analysis, !is.na(speaker_gender)), family = "binomial") 

model_compliment_s_gender_null <- glmer(compliment ~ 1 + (1|session_id/talk_id), data = subset(data_analysis, !is.na(speaker_gender)), family = "binomial") 

anova(model_compliment_s_gender, model_compliment_s_gender_null) # not significant


## speaker career stage

model_compliment_s_career <- glmer(compliment ~ speaker_career_short + (1|session_id/talk_id), data = subset(data_analysis, !is.na(speaker_career_short)), family = "binomial") 

model_compliment_s_career_null <- glmer(compliment ~ 1 + (1|session_id/talk_id), data = subset(data_analysis, !is.na(speaker_career_short)), family = "binomial") 

anova(model_compliment_s_career, model_compliment_s_career_null) # almost significant

## final model: nothing significant
```

### Giving

```{r compliment_give, echo=T, warning=F, message=F}

## question nr 

model_compliment_q_nr <- glmer(compliment ~ question_nr + (1|session_id/talk_id), data = subset(data_analysis, !is.na(question_nr)), family = "binomial") 

model_compliment_q_nr_null <- glmer(compliment ~ 1 + (1|session_id/talk_id), data = subset(data_analysis, !is.na(question_nr)), family = "binomial") 

anova(model_compliment_q_nr, model_compliment_q_nr_null) # significant

## questioner gender

model_compliment_q_gender <- glmer(compliment ~ questioner_gender + (1|session_id/talk_id), data = subset(data_analysis, !is.na(questioner_gender)), family = "binomial") 

model_compliment_q_gender_null <- glmer(compliment ~ 1 + (1|session_id/talk_id), data = subset(data_analysis, !is.na(questioner_gender)), family = "binomial") 

anova(model_compliment_q_gender, model_compliment_q_gender_null) # not significant

## questioner age

model_compliment_q_age <- glmer(compliment ~ questioner_age + (1|session_id/talk_id), data = subset(data_analysis, !is.na(questioner_age)), family = "binomial") 

model_compliment_q_age_null <- glmer(compliment ~ 1 + (1|session_id/talk_id), data = subset(data_analysis, !is.na(questioner_age)), family = "binomial") 

anova(model_compliment_q_age, model_compliment_q_age_null) # significant

# only question number and age questioner significant (LRT p < 0.05)

# build final model

model_compliment_null <- glmer(compliment ~ 1 + (1|session_id/talk_id), data = subset(data_analysis, !is.na(question_nr) & !is.na(questioner_age)), family = "binomial") 

model_compliment <- glmer(compliment ~ question_nr + questioner_age + (1|session_id/talk_id), data = subset(data_analysis, !is.na(question_nr) & !is.na(questioner_age)), family = "binomial") 

# use helper function to collect output

model_compliment_out <- collect_out(model = model_compliment, null = model_compliment_null, n_factors = 3, name = "m_compliment", type = "qa", save = "yes", dir = "../results/question-asking/")

model_compliment_out %>% t() %>% kbl() %>%
  kable_classic_2()

```

These results show is that the likelihood of a questioner giving a words of positive appraisal is higher for questions asked earlier in the Q&A, and that especially younger audience members (age \< 35 years) are more likely to give words of positive appraisal compared to older age categories. Gender of either the question asker or speaker were not significant.

## Jumpers

Next, we asked whether men or women are more likely to ask a question without being allocated to do so (i.e. chosen by the session host to ask your question). Since this might have to do with the perceived 'authority' of the session host, we controlled for the age and gender of the session host as well.

We used binomial GLMMs to address the effect of the gender and age of the questioner, and corrected for question number and age and gender of the session host. We assessed the fit of each variable using a likelihood ratio test and included only the variables that explained significant variation in the final model.

```{r jumper, echo=T, warning=F, message=F}

# recode NA to 'no jumper' = 0
jumperdata <- data_analysis %>% mutate(jumper = as.factor(case_when(
  is.na(jumper) ~ "0",
  jumper == "1" ~ "1" )))

# explore the data 
table(jumperdata$jumper, jumperdata$questioner_gender) %>% kbl() %>%
  kable_classic_2()

table(jumperdata$jumper, jumperdata$questioner_age) %>% kbl() %>%
  kable_classic_2()

# build initial models

## question nr

model_jumper_q_nr <- glmer(jumper ~ question_nr + (1|session_id/talk_id), data = subset(jumperdata, !is.na(question_nr)), family = "binomial") 

model_jumper_q_nr_null <- glmer(jumper ~ 1 + (1|session_id/talk_id), data = subset(jumperdata, !is.na(question_nr)), family = "binomial") 

anova(model_jumper_q_nr, model_jumper_q_nr_null) # not significant

## questioner age

model_jumper_q_age <- glmer(jumper ~ questioner_age + (1|session_id/talk_id), data = subset(jumperdata, !is.na(questioner_age)), family = "binomial") 

model_jumper_q_age_null <- glmer(jumper ~ 1 + (1|session_id/talk_id), data = subset(jumperdata, !is.na(questioner_age)), family = "binomial") 

anova(model_jumper_q_age, model_jumper_q_age_null) # not significant

## questioner gender

model_jumper_q_gender <- glmer(jumper ~ questioner_gender + (1|session_id/talk_id), data = subset(jumperdata, !is.na(questioner_gender)), family = "binomial") 

model_jumper_q_gender_null <- glmer(jumper ~ 1 + (1|session_id/talk_id), data = subset(jumperdata, !is.na(questioner_gender)), family = "binomial") 

anova(model_jumper_q_gender, model_jumper_q_gender_null) # almost significant

summary(model_jumper_q_gender)

## host age

model_jumper_h_age <- glmer(jumper ~ host_1_age + (1|session_id/talk_id), data = subset(jumperdata, !is.na(host_1_age)), family = "binomial") 

model_jumper_h_age_null <- glmer(jumper ~ 1 + (1|session_id/talk_id), data = subset(jumperdata, !is.na(host_1_age)), family = "binomial") 

anova(model_jumper_h_age, model_jumper_h_age_null) # not significant

## host gender
 
model_jumper_h_gender <- glmer(jumper ~ host_1_gender + (1|session_id/talk_id), data = subset(jumperdata, !is.na(host_1_gender)), family = "binomial") 

model_jumper_h_gender_null <- glmer(jumper ~ 1 + (1|session_id/talk_id), data = subset(jumperdata, !is.na(host_1_gender)), family = "binomial") 

anova(model_jumper_h_gender, model_jumper_h_gender_null) # not significant

# nothing is significant, trend for gender questioner


```

These results show is that the likelihood of a person jumping a question not significantly affected by any of the variables. There was however a tendency for men to be more likely to jump a question compared to women. Note that the inference of these models is however limited, since jumpers were rare (N = `r nrow(subset(data_analysis, jumper ==1))`).

## Speaker over time

Next, we investigated the probability that a speaker talks for longer than their allocated speaking time is affected by speaker gender or career stage. We did not expect that any other confounding variables would explain variation in speaking overtime, since this is something that was prepared by only the speaker.

```{r overtime, echo=T, warning=F, message=F}

# have to summarize by talk not per question
data_overtime <- data_analysis %>% select(talk_id, session_id, overtime, speaker_gender, speaker_career_short) %>% unique()

# N = 
nrow(data_overtime)

# explore data
table(data_overtime$overtime, data_overtime$speaker_gender) %>% kbl() %>%
  kable_classic_2()

table(data_overtime$overtime, data_overtime$speaker_career_short) %>% kbl() %>%
  kable_classic_2()
# build initial models

## speaker gender
model_overtime_s_gender <- glmer(overtime ~ speaker_gender + (1|session_id/talk_id), data = subset(data_overtime, !is.na(speaker_gender)), family = "binomial") 

model_overtime_s_gender_null <- glmer(overtime ~ 1 + (1|session_id/talk_id), data = subset(data_overtime, !is.na(speaker_gender)), family = "binomial") 

anova(model_overtime_s_gender, model_overtime_s_gender_null) # not significant

## speaker career 
model_overtime_s_career <- glmer(overtime ~ speaker_career_short + (1|session_id/talk_id), data = subset(data_overtime, !is.na(speaker_career_short)), family = "binomial") 

model_overtime_s_career_null <- glmer(overtime ~ 1 + (1|session_id/talk_id), data = subset(data_overtime, !is.na(speaker_career_short)), family = "binomial") 

anova(model_overtime_s_career, model_overtime_s_career_null) # significant

summary(model_overtime_s_career)

# use helper function to collect output

model_overtime_out <- collect_out(model = model_overtime_s_career, null = model_overtime_s_career_null, n_factors = 2, name = "m_overtime", type = "qa", save = "yes", dir = "../results/question-asking/")

model_overtime_out %>% t() %>% kbl() %>%
  kable_classic_2()

```

The model output shows that late career speakers were more likely to speak overtime and mid-career speakers were the least likely to speak overtime.

## Critical question

Next, we assessed whether the likelihood of receiving a 'critical' question is affected by the gender and age of the questioner or the gender and career stage of the speaker.

### Receiving

```{r critical_receive, echo=T, warning=F, message=F}

# explore data
table(data_analysis$question_type_e, data_analysis$questioner_gender) %>% kbl() %>%  kable_classic_2()

table(data_analysis$question_type_e, data_analysis$questioner_age) %>% kbl() %>%  kable_classic_2()

table(data_analysis$question_type_e, data_analysis$speaker_gender) %>% kbl() %>%
  kable_classic_2()

table(data_analysis$question_type_e, data_analysis$speaker_career_short) %>% kbl() %>%  kable_classic_2()

# build initial models

## speaker gender
model_critical_s_gender <- glmer(question_type_e ~ speaker_gender + (1|session_id/talk_id), data = subset(data_analysis, !is.na(speaker_gender)), family = "binomial") 

model_critical_s_gender_null <- glmer(question_type_e ~ 1 + (1|session_id/talk_id), data = subset(data_analysis, !is.na(speaker_gender)), family = "binomial") 

anova(model_critical_s_gender, model_critical_s_gender_null) # not significant, sort of trend

## speaker career stage
model_critical_s_career <- glmer(question_type_e ~ speaker_career_short + (1|session_id/talk_id), data = subset(data_analysis, !is.na(speaker_career_short)), family = "binomial") 

model_critical_s_career_null <- glmer(question_type_e ~ 1 + (1|session_id/talk_id), data = subset(data_analysis, !is.na(speaker_career_short)), family = "binomial") 

anova(model_critical_s_career, model_critical_s_career_null) # not significant

## nothing significant
```

### Giving

```{r critical_give, echo=T, warning=F, message=F}
## questioner gender
model_critical_q_gender <- glmer(question_type_e ~ questioner_gender + (1|session_id/talk_id), data = subset(data_analysis, !is.na(questioner_gender)), family = "binomial") 

model_critical_q_gender_null <- glmer(question_type_e ~ 1 + (1|session_id/talk_id), data = subset(data_analysis, !is.na(questioner_gender)), family = "binomial") 

anova(model_critical_q_gender, model_critical_q_gender_null) # not significant

## questioner age
model_critical_q_age <- glmer(question_type_e ~ questioner_age + (1|session_id/talk_id), data = subset(data_analysis, !is.na(questioner_age)), family = "binomial") 

model_critical_q_age_null <- glmer(question_type_e ~ 1 + (1|session_id/talk_id), data = subset(data_analysis, !is.na(questioner_age)), family = "binomial") 

anova(model_critical_q_age, model_critical_q_age_null) # significant, older higher probability

model_critical_q_age_out <- collect_out(model = model_critical_q_age, null = model_critical_q_age_null, n_factors = 2, name = "m_critical_q_age", type = "qa", save = "yes", dir = "../results/question-asking/")

model_critical_q_age_out %>% t() %>% kbl() %>%  kable_classic_2()

# only age questioner significant
```

These results show is that the likelihood of a person asking or receiving a critical question was affected by the age category of the question asker. Note that the inference of these models is however limited, since critical questions were rare and subject to observer bias (N = `r nrow(subset(data_analysis, question_type_e ==1))`).
